Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2017-04-13 manuscript.tex
latest

By contrast, models of phenotypic trait evolution represent the change in phenotypic abundances over time using absolute fitness functions which describe how those traits affect survival and reproduction in particular ecological scenarios. This approach is powerful enough to model eco-evolutionary feedbacks between co-evolving traits, but is generally problem-specific and restricted to only a few traits at a time. These models emphasize invasion from low frequencies and co-existence more than frequency trajectories over time, with the entire adaptive dynamics approach revolving around ``invasion fitness'' \citep{diekmann_2004}.

@@

Nevertheless,  broader principles governing the operation of selection.

In many groups of organisms, including corals \citep{darling_2012}, insects \citep{southwood_1977}, fishes \citep{winemiller_1992}, zooplankton \citep{allan_76} and plants \citep{grime_1988,westoby_1998},

@@

The most famous example is Grime's plant trait classification scheme \citep{grime_1974,grime_1977,grime_1988}. Grime considered two broad determinants of population density: stress (persistent hardship e.g. due to resource scarcity, unfavorable temperatures or toxins) and disturbance (intermittent destruction of vegetation e.g. due to trampling, herbivory, pathogens, extreme weather or fire).  The extremes of these two factors define three primary strategies denoted by C/S/R respectively (Fig. \ref{fig:grimeschematic}): competitors ``C'' excel in low stress, low disturbance environments; stress tolerators ``S'' excel in high stress, low disturbance environments; and ruderals ``R''  excel in low stress, high disturbance environments. Survival is not possible in high-stress, high-disturbance environments. Grime showed that measures of C, S and R across a wide range of plant species are anti-correlated, so that strong C-strategists are weak S and R strategists, and so on. Thus, plant species can be classified on a triangular C/S/R ternary plot \citep{grime_1974}. Trait classification schemes for other organisms are broadly analogous to Grime's scheme \citep{winemiller_2015}.

>>>>

By contrast, models of phenotypic trait evolution represent the change in phenotypic abundances over time using absolute fitness functions which describe how those traits affect survival and reproduction in particular ecological scenarios. This approach is powerful enough to model eco-evolutionary feedbacks between co-evolving traits, but is generally problem-specific and restricted to only a few traits at a time. These models emphasize invasion from low frequencies and co-existence over frequency trajectories over time. For instance, ``invasion fitness'' is the foundation of the adaptive dynamics approach \citep{diekmann_2004}.

@@

Empirical studies of organismal traits corals \citep{darling_2012}, insects \citep{southwood_1977}, fishes \citep{winemiller_1992}, zooplankton \citep{allan_76} and plants \citep{grime_1988,westoby_1998},

@@

The most famous example is Grime's plant trait classification scheme \citep{grime_1974,grime_1977,grime_1988}. Grime considered two broad determinants of population density: stress (persistent hardship e.g. due to resource scarcity, unfavorable temperatures or toxins) and disturbance (intermittent destruction of vegetation e.g. due to trampling, herbivory, pathogens, extreme weather or fire).  The extremes of these two factors define three primary strategies denoted by C/S/R respectively (Fig. \ref{fig:grimeschematic}): competitors ``C'' excel in low stress, low disturbance environments; stress tolerators ``S'' excel in high stress, low disturbance environments; and ruderals ``R''  excel in low stress, high disturbance environments. Survival is not possible in high-stress, high-disturbance environments. Grime showed that measures of C, S and R across a wide range of plant species are anti-correlated, so that strong C-strategists are weak S and R strategists, and so on. Thus, plant species can be classified on a triangular C/S/R ternary plot \citep{grime_1974}.

Trait classification schemes for other organisms are broadly analogous to Grime's scheme \citep{winemiller_2015}.
